---
title: "Statistical Inference Assignment Project Part 2"
author: "Francisco Mimica Porras"
date: "2/16/2021"
output:
  word_document: default
  html_document: default
---


# Overview

This report aims to analyze the ToothGrowth data in the R datasets package. Per the course project instructions, the following items should occur:

Load the ToothGrowth data and perform some basic exploratory data analyses
Provide a basic summary of the data.
Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose (only use the techniques from class, even if there’s other approaches worth considering).
State your conclusions and the assumptions needed for your conclusions.`

# Analysis

# First we will load the necessary libraries.

```{r}
library(ggplot2)

```

# Load ToothGrowth data

```{r}
data("ToothGrowth")
```

# Display a summary of the data

```{r}
summary(ToothGrowth)
```
#Display the first few rows of data

```{r}
head(ToothGrowth)
```

##Unique Values
```{r}
unique(ToothGrowth$len)
unique(ToothGrowth$supp)
unique(ToothGrowth$dose)
```

# Plots

# exploring the data.
## Convert dose to a factor
```{r}
ToothGrowth$dose<-as.factor(ToothGrowth$dose)
```

## Plot tooth length ('len') vs. the dose amount ('dose'), broken out by supplement delivery method ('supp')

```{r}
ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose)) + xlab("Dose Amount") + ylab("Tooth Length") + facet_grid(~ supp) + ggtitle("Tooth Length vs. Dose Amount \nby Delivery Method") + 
     theme(plot.title = element_text(lineheight=.8, face="bold"))
```

## Plot tooth length ('len') vs. supplement delivery method ('supp') broken out by the dose amount ('dose')

```{r}
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp)) + xlab("Supplement Delivery") + ylab("Tooth Length") + facet_grid(~ dose) + ggtitle("Tooth Length vs. Delivery Method \nby Dose Amount") + 
     theme(plot.title = element_text(lineheight=.8, face="bold"))
```

# Comparing tooth growth by supplement using a t-test.

# run t-test

```{r}
t.test(len~supp,data=ToothGrowth)
```

### The p-value of this test was 0.06. Since the p-value is greater than 0.05 and the confidence interval of the test contains zero we can say that supplement types seems to have no impact on Tooth growth based on this test.


# Comparing tooth growth by dose, looking at the different pairs of dose values.

## Run t-test using dose amounts 0.5 and 1.0
```{r}
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,0.5))
t.test(len~dose,data=ToothGrowth_sub)
```

## Run t-test using dose amounts 0.5 and 2.0

```{r}
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(0.5,2.0))
t.test(len~dose,data=ToothGrowth_sub)
```

## Run t-test using dose amounts 1.0 and 2.0

```{r}
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,2.0))
t.test(len~dose,data=ToothGrowth_sub)
```

The p-value of each test was essentially zero and the confidence interval of each test does not cross over zero (0). We can assume that the average tooth length increases with an inceasing dose, and therefore the null hypothesis can be rejected.

# Conclusion

The t-test analysis from above, we can conclude that supplement delivery has no effect on tooth growth/length, however increased dosages do result in increased tooth length.

























